---
title: "Demonstration of `itses`"
author: "Amir Ahmed"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This `vignette` provides a quick introducion to the problem treated by the `itses` package and a demonstration of how to use the package in practice.

## Introduction

Let $\mathbf{Y} = (Y_1, Y_2, \dots, Y_n)$. Assume $Y_i \sim N(\theta_i, \sigma^2)$,  $\theta_i$ unknown and $\sigma^2>0$ known and given or unknown, for all $1\leq i\leq n$. That is $Y_i$ of normal distribuion with mean $\theta_i$ and variance $\sigma^2$. Estimation is to be done minimizing expected square error under repeated sampling. 

`itses` estimate the means $\theta_i$ using iterative sparse estimation. Risk estimation is either done using sampling or numerically.

Noise estimation is done using Sparse Median Absolute Deviation or by Median Absolute Deviation.
## A basic example

First fetch the package and simulate some data:

```{r, fig.show='hold'}
library(itses)
set.seed(1)
# Simulate some data to be used
theta <- c(rep(0, 3), 2*rnorm(7))
theta <- sample(theta)
y <- rnorm(length(theta), mean = theta, sd = 3)
```

The true means are:
````{r}
theta

```` 

The observed data is: 
````{r}
y

```` 

Iterative sparse estimation with the soft-threshold estimator for the means and Sparse MAD noise estimation can be done using:
````{r}
theta.est.st.sd.unknown <- itses::itses(y)$theta
theta.est.st.sd.unknown
```` 

If noise is known it can be provided by:
````{r}
theta.est.st.sd.known <- itses::itses(y, sd = 3)$theta
theta.est.st.sd.known
```` 

To use hard-threshold one can instead run:
````{r}
theta.est.ht.sd.known <- itses::itses(y, sd = 3, method = "HT")$theta
theta.est.ht.sd.known 
```` 


## Plotting estimated risk 
Estimated risk of final iteration `itses` can be plotted in the following manner: 
 
````{r}
itses.data <- itses::itses(y, sd = 3, method = "ST")
iteration.data <- itses.data$iteration_result
last.iteration.data <- iteration.data[[length(iteration.data)]]
```` 

 
````{r}
plot(last.iteration.data[,1], last.iteration.data[,2], type = "l",
     xlab = "threshold", ylab = "Risk (squared error)")
```` 

To display estimated risk by all iterations one can similarly do:
```{r}
for(i in length(iteration.data):1) {
  if(i == length(iteration.data)) {
    plot(iteration.data[[i]][,1], iteration.data[[i]][,2], type = "l",
     xlab = "threshold", ylab = "Risk (squared error)")
  }else{
    lines(iteration.data[[i]][,1], iteration.data[[i]][,2], col = "gray")
  }
}
```
Similarly estiamted risk at each iteration threshold can be plotted with:
```{r}

y <- c(rnorm(30), rnorm(30)+10) # Just generating some data
itses.data <- itses::itses(y, sd = 3, method = "ST")
iteration.data <- itses.data$iteration_result
risks <- c()
for(i in 1:length(iteration.data)) {
  data <- iteration.data[[i]]
  min.risk <- min(data[,2])
  risks <- c(risks, min.risk)
}
plot(1:length(risks), risks, xlab = "iteration", ylab = "estimated minimum risk", type = "l")
```

## Estimating noise

To only estimate noise the SparseMAD noise estimator has been made available. To run:
